Concurrent Programming in Harmony

### 2022-01-06 book/paper.tex
Latest

Harmony is heavily influenced by Leslie Lamport's work on
TLA+, TLC, and PlusCal~\cite{Lamport02, Lamport09},
Gerard Holzmann's work on Promela and SPIN~\cite{SPIN},
and University of Washington's DSLabs system~\cite{MWA19}.
Harmony is designed to have a lower learning curve than those
systems, but is not as powerful.  When you finish this book
and want to learn more, we strongly encourage checking
those out.
Another excellent resource is Fred Schneider's book ``On
Concurrent Programming''~\cite{Schneider97}.
(This chapter is named after that book.)

@@

The appendices contain various details about Harmony itself, including
an appendix that explains how Harmony works (\autoref{app:howitworks}) and
an appendix on installing and running Harmony (\autoref{app:install}).

@@

\begin{itemize}
\item <{False < True}>;
\item integers are ordered in the natural way;
\item strings are lexicographically ordered;
\item program counters are ordered by their integer values;
\item dictionaries are first converted into a list of ordered (key, value)
pairs.  Then two dictionaries are lexicographically ordered by this
representation;
\item a set is first converted into an ordered list, then lexicographically
ordered;
\item an address is a list of Harmony values.  <{None}> is the empty list.
Addresses are lexicographically ordered accordingly;
\item contexts (\autoref{app:context}) are ordered first by name, then by program counter, then by the remaining content in an unspecified way.
\end{itemize}

@@

\chapter{The Harmony Virtual Machine}\label{app:hvm}

@@

\section{Contexts and Threads}
\label{app:context}

@@

\chapter{How Harmony Works}\label{app:howitworks}

@@

\begin{enumerate}
\item The Harmony \emph{compiler} turns your Harmony program into
bytecode.  A recursive descent parser and code generator written in
Python (see \texttt{harmony.py}) turns an \texttt{x.hny} program
into \texttt{x.hvm}, a JSON file containing the corresponding
bytecode.
\item The Harmony \emph{model checker} evaluates the state space
that the program (now in bytecode) can generate.  The model checker
is written in C as it needs to be highly efficient (see \texttt{charm.c}).
The model checker starts from the initial state, and then, iteratively,
checks for each state that it has found what next steps are possible and
generates the next states using the Harmony virtual machine
(\autoref{app:hvm}).  If the model is finite, eventually the model
checker will generate a graph with all possible states.
If there is a problematic path in this graph (see below), then it will
report the shortest such path in the \texttt{x.hco} output file in JSON
format.
\item The \texttt{x.hco} output file is translated twice by
\texttt{harmony.py}.  There is a so-called \emph{brief output} that
is written to standard output.  The rest depends on whether there was
a problem with the execution or not.  If there was a problem, the more
comprehensive output is placed in the \texttt{x.htm} HTML output file,
allowing you to navigate the problematic path and all the details of
each of the states on the path.  If not, a DFA of the print behavior
is generated and compared with a provided DFA if specified with the
\texttt{-B} flag.
\end{enumerate}

@@

\chapter{Installing and Running Harmony}\label{app:install}

>>>>

Harmony is heavily influenced by Leslie Lamport's work on
TLA+, TLC, and PlusCal~\cite{Lamport02, Lamport09},
Gerard Holzmann's work on Promela and SPIN~\cite{SPIN},
and University of Washington's DSLabs system~\cite{MWA19}.
Some of the examples in this book are derived from those sources.
Harmony is designed to have a lower learning curve than those
systems, but is not as powerful.  When you finish this book
and want to learn more, we strongly encourage checking
those out.
Another excellent resource is Fred Schneider's book ``On
Concurrent Programming''~\cite{Schneider97}.
(This chapter is named after that book.)

@@

If you already know about concurrent and distributed programming and
are just interested in a ``speed course'' on Harmony, I would recommend
reading \autoref{ch:harmonyintro}, \autoref{ch:harmonymachine},
\autoref{ch:method}, \autoref{ch:specification}, and \autoref{ch:cds}.
The appendices contain various details about Harmony itself, including
an appendix on convenient Harmony modules (\autoref{ap:module}),
an appendix that explains how Harmony works (\autoref{ap:howitworks}),
and
an appendix on installing and running Harmony (\autoref{ap:install}).

@@

\begin{itemize}
\item <{False < True}>;
\item integers are ordered in the natural way;
\item strings are lexicographically ordered;
\item program counters are ordered by their integer values;
\item dictionaries are first converted into a list of ordered (key, value)
pairs.  Then two dictionaries are lexicographically ordered by this
representation;
\item a set is first converted into an ordered list, then lexicographically
ordered;
\item an address is a list of Harmony values.  <{None}> is the empty list.
Addresses are lexicographically ordered accordingly;
\item contexts (\autoref{ap:context}) are ordered first by name, then by program counter, then by the remaining content in an unspecified way.
\end{itemize}

@@

\chapter{The Harmony Virtual Machine}\label{ap:hvm}

@@

\section{Contexts and Threads}
\label{ap:context}

@@

\chapter{How Harmony Works}\label{ap:howitworks}

@@

\begin{enumerate}
\item The Harmony \emph{compiler} turns your Harmony program into
bytecode.  A recursive descent parser and code generator written in
Python (see \texttt{harmony.py}) turns an \texttt{x.hny} program
into \texttt{x.hvm}, a JSON file containing the corresponding
bytecode.
\item The Harmony \emph{model checker} evaluates the state space
that the program (now in bytecode) can generate.  The model checker
is written in C as it needs to be highly efficient (see \texttt{charm.c}).
The model checker starts from the initial state, and then, iteratively,
checks for each state that it has found what next steps are possible and
generates the next states using the Harmony virtual machine
(\autoref{ap:hvm}).  If the model is finite, eventually the model
checker will generate a graph with all possible states.
If there is a problematic path in this graph (see below), then it will
report the shortest such path in the \texttt{x.hco} output file in JSON
format.
\item The \texttt{x.hco} output file is translated twice by
\texttt{harmony.py}.  There is a so-called \emph{brief output} that
is written to standard output.  The rest depends on whether there was
a problem with the execution or not.  If there was a problem, the more
comprehensive output is placed in the \texttt{x.htm} HTML output file,
allowing you to navigate the problematic path and all the details of
each of the states on the path.  If not, a DFA of the print behavior
is generated and compared with a provided DFA if specified with the
\texttt{-B} flag.
\end{enumerate}

@@

\chapter{Installing and Running Harmony}\label{ap:install}
